Switch equivalence and weak prime domains for fusions

### 2024-08-18 fusione.tex
reticoli dist

 \begin{theorem}[\cite{garner2012axioms,castelnovo2023thesis,johnstone2007quasitoposes}]label{prop:un}Let $\X$ be a strict $\mathcal{M}$-adhesive category with pullbacks. If $\mathcal{M}$ contains all split monomorphisms  then for every onject $X$, $(\sub{M}{X}{X}, \leq)$ has binary suprema.
 \end{theorem}
\begin{proof}	Let $m\colon M\to X$  and $n\colon N\to X$ be elements of in $\mathcal{M}$. By \Cref{prop:uni2} we know that $[m]$ and $[n]$  have a supremum in $(\msub{X}{X}, \leq)$ given by $[u]$, where $u\colon U\to X$ fits in the  diagram
	\[\xymatrix{P \ar@{>->}[r]^{p_1} \ar@{>->}[d]_{p_2} & M\ar@{>->}[d]^{u_2} \ar@{>->}@/^.3cm/[ddr]^{m} \\ N \ar@{>->}@/_.3cm/[drr]_{n}\ar@{>->}[r]_{u_1} & U \ar[dr]^{u} \\ && X}\]
	in which the outer edges form a pullback and the inner square is a pushout.

@@

 \begin{corollary}
 	\todo{reticolo distributivo}
 \end{corollary}
\begin{proof}
	contenuto...
\end{proof}

\begin{example}\label{ex:sem}
	\todo{nei grafi semplici questa cosa non funziona}
\end{example}

>>>>

 \begin{theorem}[\cite{garner2012axioms,castelnovo2023thesis,johnstone2007quasitoposes}]\label{thm:un}Let $\X$ be a strict $\mathcal{M}$-adhesive category with pullbacks and suppose that $\mathcal{M}$ contains all split monomorphisms. Then for every object $X$, $(\sub{M}{X}{X}, \leq)$ has binary suprema, which are preserved by the inclusion $(\sub{M}{X}{X}, \leq)\to (\msub{X}{X}, \leq)$.
 \end{theorem}
\begin{proof}	Let $m\colon M\to X$  and $n\colon N\to X$ be elements of in $\mathcal{M}$. By \Cref{prop:uni2} we know that $[m]$ and $[n]$  have a supremum in $(\msub{X}{X}, \leq)$ given by $[u]$, where $u\colon U\to X$ fits in the  diagram
	\[\xymatrix{P \ar@{>->}[r]^{p_1} \ar@{>->}[d]_{p_2} & M\ar@{>->}[d]^{u_2} \ar@{>->}@/^.3cm/[ddr]^{m} \\ N \ar@{>->}@/_.3cm/[drr]_{n}\ar@{>->}[r]_{u_1} & U \ar[dr]^{u} \\ && X}\]
	in which the outer edges form a pullback and the inner square is a pushout.

@@

\begin{example}\label{ex:sem}In \cite{johnstone2007quasitoposes} it is shown that the category $\cat{SGraph}$ of simple graphs is not quasiadhesive. Too see this, first of all recall that a morphism in $\cat{SGraph}$ is a regular monos if and only if it is injective on vertex and reflects edges. Now, take 
	\[\mathcal{G}:=(1, \{0,1\}, \delta_0, \delta_1) \qquad \mathcal{H}:=( \emptyset, 1, ?_1, ?_1) \quad \]
	Then we have two regular monomorphisms $ (?_1, \delta_0), (?_1, \delta_1)\colon \mathcal{H}\rightrightarrows \mathcal{G} $. Their supremum in $\msub{SGraph}{\mathcal{G}}$ is represented by inclusion of $(\emptyset, \{0,1\}, ?_0, ?_1)$ into $\mathcal{G}$ which is not a regular monomorphism, thus, by \Cref{thm:un}, $\cat{SGraph}$ is not quasiadhesive. 
\end{example}

 \begin{proposition}\label{prop:distlat}Let $\X$ be a strict $\mathcal{M}$-adhesive with pullbacks. If $\mathcal{M}$ contains all split monos then for every object $X$, $(\sub{M}{X}{X})$ is a distributive lattice. 
 \end{proposition}
\begin{proof}
	Since $\X$ has pullbacks then $(\sub{M}{X}{X})$  has binary infima. The existence of binary suprema is guaranteed by \Cref{thm:un}. We only have to show that, given $m\colon M\mto X$, $n\colon N\mto X$ and $v\colon V\mto X$ we have
	\[[n]\wedge ([m]\vee[v])=([n]\wedge [m]) \vee ([n]\wedge [v])\]

	Let $u\colon \mto X$ be a representative of $[m]\vee [v]$, thus by \Cref{prop:uni2} and \Cref{thm:un} we know that it fits in the diagram below, whose outer boundary is a pullback and in which the inner square is a pushout.
		\[\xymatrix{P \ar@{>->}[r]^{p_1} \ar@{>->}[d]_{p_2} & M\ar@{>->}[d]^{u_2} \ar@{>->}@/^.3cm/[ddr]^{m} \\ V \ar@{>->}@/_.3cm/[drr]_{v}\ar@{>->}[r]_{u_1} & U \ar@{>->}[dr]^{u} \\ && X}\]

	Let also $r\colon R\mto X$,  $s\colon S\mto X$, $t\colon T\mto X$ and $q\colon Q\mto X$ be  representative for, respectively $[n]\wedge [u]$, $[n]  \wedge [m]$, $[n]\wedge [v]$ and $[s]\wedge [t]$,  then we have the four pullback squares below.
	\[\xymatrix{R \ar@{>->}[r]^{r_1} \ar@{>->}[d]_{r_2}\ar@{>->}[dr]^{r}& N \ar@{>->}[d]^{n} & S \ar@{>->}[r]^{s_1} \ar@{>->}[d]_{s_2}\ar@{>->}[dr]^{s}& N \ar@{>->}[d]^{n}&T \ar@{>->}[r]^{t_1} \ar@{>->}[d]_{t_2}\ar@{>->}[dr]^{t}& N \ar@{>->}[d]^{n}& Q \ar@{>->}[r]^{q_1} \ar@{>->}[d]_{q_2}\ar@{>->}[dr]^{q}& T \ar@{>->}[d]^{t}\\U \ar@{>->}[r]_{u} & X&M \ar@{>->}[r]_{m} & X&V \ar@{>->}[r]_{v} & X&S \ar@{>->}[r]_{s} & X}\]

Using the universal property of the first square, we can construct the two dotted arrows $m_1\colon T\to R$ and $m_2\colon S\to R$ fitting in the following diagrams. Notice that the left halves of these rectangles are pullbacks by \Cref{lem:pb1}.

\[\xymatrix{T \ar[d]_{t_2} \ar@/^.4cm/[rr]^{t_1} \ar@{.>}[r]_{m_1}& R \ar@{>->}[r]_{r_1} \ar@{>->}[d]_{r_2}& N \ar@{>->}[d]^{n} & S \ar[d]_{s_2} \ar@/^.4cm/[rr]^{s_1} \ar@{.>}[r]_{m_2}& R \ar@{>->}[r]_{r_1} \ar@{>->}[d]_{r_2}& N \ar@{>->}[d]^{n}\\ V \ar@/_.4cm/[rr]_{v} \ar@{>->}[r]^{u_1}&U \ar@{>->}[r]^{u} & X &  M \ar@/_.4cm/[rr]_{m} \ar@{>->}[r]^{u_2}&U \ar@{>->}[r]^{u} & X}\]

On the other hand, we can also build the dotted $j\colon Q\to P$ as follows.

		\[\xymatrix{Q \ar[r]^{q_2} \ar[d]_{q_1} \ar@{.>}[dr]^{j}& S \ar@{>->}@/^.2cm/[dr]^{s_2}\\ T \ar@{>->}@/_.2cm/[dr]_{t_2}&P \ar@{>->}[r]^{p_1} \ar@{>->}[d]_{p_2} & M\ar@{>->}[d]^{u_2}  \\ &V \ar@{>->}[r]_{u_1} & U }\]

	Thus we have the commutative cube below, whose front and right faces are pullbacks and whose bottom face is an $\mathcal{M}$-pushout.
		\[\xymatrix@C=15pt@R=15pt{&Q\ar[dd]|\hole_(.65){j}\ar@{>->}[rr]^{q_2} \ar@{>->}[dl]_{q_1} && S \ar@{>->}[dd]^{s_2} \ar@{>->}[dl]_{m_2}\\ T \ar@{>->}[dd]_{t_2}\ar@{>->}[rr]^(.65){m_1} & & R \ar@{>->}[dd]_(.3){r_2}\\&P\ar@{>->}[rr]|\hole^(.7){p_1} \ar@{>->}[dl]_{p_2} && M \ar@{>->}[dl]^{u_2} \\V \ar@{>->}[rr]_{u_1} & & U}\]

		To show that the top face is a pushout it is enough to show that the left and back faces are pullbacks. Suppose thus that the solid part of the following two diagrams are given.
		\[\xymatrix{Z \ar@{.>}[dr]^{z} \ar@/^.3cm/[drr]^{z_1} \ar@/_.3cm/[ddr]_{z_2}&& &Y \ar@{.>}[dr]^{y} \ar@/^.3cm/[drr]^{y_1} \ar@/_.3cm/[ddr]_{y_2}\\&Q \ar@{>->}[r]^{q_2} \ar[d]_j & S \ar@{>->}[d]^{s_2}& &Q \ar@{>->}[r]^{q_1} \ar[d]_{j} & T \ar@{>->}[d]^{t_2}\\& P \ar@{>->}[r]_{p_1} & M && P \ar@{>->}[r]_{p_2}& V}\]

		By the universal property of pullbacks, we get the dotted arrows $z'\colon Z\to T$ and $y'\colon Y\to S$ fitting in the diagrams below.
		\[\xymatrix{Z \ar@{.>}[dr]^{z'} \ar[r]^{z_1} \ar[d]_{z_2}& S \ar@{>->}@/^.2cm/[dr]^{s_1}& &Y \ar@{.>}[dr]^{y'} \ar[r]^{y_1} \ar[d]_{y_2} & T  \ar@{>->}@/^.2cm/[dr]^{t_1}\\P \ar@{>->}@/_.2cm/[dr]_{p_2}&T \ar@{>->}[r]^{t_1} \ar@{>->}[d]_{t_2} & N \ar@{>->}[d]^{s_2}& P\ar@{>->}@/_.2cm/[dr]_{p_1}&S \ar@{>->}[r]^{s_1} \ar@{>->}[d]_{s_2} & N \ar@{>->}[d]^{n}\\& V \ar@{>->}[r]_{v} & X && M \ar@{>->}[r]_{m}& X}\]

		In turn, these two arrows yields $\hat{z}\colon Z \to Q$ and $\hat{y}\colon Y \to Q$ as below.
				\[\xymatrix{Z \ar@{.>}[dr]^{\hat{z}} \ar@/^.3cm/[drr]^{z_1} \ar@/_.3cm/[ddr]_{z'}&& &Y \ar@{.>}[dr]^{y} \ar@/^.3cm/[drr]^{y_1} \ar@/_.3cm/[ddr]_{y'}\\&Q \ar@{>->}[r]^{q_2} \ar@{>->}[d]_{q_1} & S \ar@{>->}[d]^{s}& &Q \ar@{>->}[r]^{q_1} \ar@{>->}[d]_{q_2} & T \ar@{>->}[d]^{t}\\& T \ar@{>->}[r]_{t} & X && S \ar@{>->}[r]_{s}& X}\]

	If we compute we get:
	\[\begin{split}
		p_2\circ j\circ \hat{z}&= t_2\circ q_1\circ \hat{z}\\&=t_2\circ z'\\&= p_2\circ z_2
	\end{split}\qquad \begin{split}
	p_1\circ j \circ \hat{y}&=s_2\circ q_2\circ \hat{y}\\&=s_2\circ y'\\&=p_1\circ y_2
	\end{split}\]

	Since $p_1$ and $p_2$ are mono it follows that 
	\[j\circ \hat{z}=z_2 \qquad j\circ \hat{y}=y_2\]

	Moreover, $q_2$ and $q_1$  are monos too, thus $\hat{z}$ and $\hat{y}$ are the unique arrows fitting in the two diagram with which we have begun, proving that the back and right faces of our original cubes are pullbacks.

	To end the proof it is now enough to see that 
	\[\begin{split}
	r\circ m_1&=u\circ r_2\circ m_1\\&=u\circ u_1\circ t_2\\&=v\circ t_2\\&=t
	\end{split}\qquad \begin{split}
	r\circ m_2&=u\circ r_2\circ m_2\\&=u\circ u_2\circ s_2\\&=m\circ s_2\\&=s
	\end{split}\]

Thus by \Cref{prop:uni2,thm:un} $[r]$ is the supremum of $[s]$ and $[t]$, which is exactly the thesis.	
\end{proof}
